Formal Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2022-11-04 nakayama/paper.tex
Edits to prose and math explanation

The set of integers also forms a \emph{ring} due to its properties of multiplication.  
It is closed under this binary operation, which satisfies associativity.  
The integer $1$ is a   \emph{multiplicative identity} since for any  $n \in \mathbb{Z}$ one has $n \cdot 1 = 1 \cdot n = n$.  Finally, the the distributive law governing the compatibility of addition and multiplication holds. 
Even more, the integers form a 
\emph{commutative ring} since $n \cdot m = m \cdot n$ for all integers $n$ and $m$.

@@

The ring of polynomials over a field $F$ in a variable $x$ is not local; in fact, given any irreducible polynomial $f(x)$, the set of its multiples is a maximal ideal of the polynomial ring $F[x]$.
On the other hand , the ring of formal power series over $F$ in $x$  is local; its maximal ideal consists of the power series with no constant term.

@@

We started by defining a semigroup class, which declares a binary operation to
be associative. From here, we build up through monoids, which introduce
identities, to groups, which introduce inverses. Note the double equals in
these definitions is notation for an arbitrary equivalence relation over the
group's carrier set which acts as equality.
% \begin{verbatim}
\begin{minted}{coq}
  Infix "==" := equiv (at level 60, no associativity).
  Class Semigroup := {
    semigroup_assoc:
      forall (a b c: Carrier),
        a <o> b <o> c == a <o> (b <o> c);
  }.
  Class Monoid := {
    monoid_semigroup :> Semigroup equiv op;
    monoid_ident_l:
      forall (a: Carrier), ident <o> a == a;
    monoid_ident_r:
      forall (a: Carrier), a <o> ident == a;
  }.
  Class Group := {
    group_monoid :> Monoid equiv op ident;
    group_inv_l:
      forall (a: Carrier), inv a <o> a == ident;
    group_inv_r:
      forall (a: Carrier), a <o> inv a == ident;
  }.
\end{minted}
% \end{verbatim}
The lines like \verb|monoid_semigroup :> Semigroup equiv op;| simply coerce
the monoid typeclass into a semigroup.

While we found later on that we did not need quotients, it is worth remarking
that we got quotients working rather nicely in Coq using typeclasses. An
algebraic quotient are an equivalence relation on the algebraic structure which
preserve that structure. For example, quotient groups are formed by taking a
subgroup and making every element of that subgroup equivalent to the identity.
With \texttt{P} being the predicate for the subgroup, there are two ways to
make an equivalence relation from this description.
% \begin{verbatim}
\begin{minted}{coq}
  Definition left_congru (a b: Carrier) :=
    P (inv a <o> b).
  Definition right_congru (a b: Carrier) :=
    P (a <o> inv b).
\end{minted}
% \end{verbatim}
When these two relations coincide, then we can prove that the equivalence
relation(s) actually preserve the group structure. Subgroups which have this
property are called \emph{normal subgroups}.
% \begin{verbatim}
\begin{minted}{coq}
  Let normal_subgroup_congru_coincide :=
    forall (a b: Carrier),
      left_congru op inv P a b <->
      right_congru op inv P a b.

  Theorem quotient_normal_subgroup_group:
    normal_subgroup_congru_coincide ->
    Group (left_congru op inv P) op ident inv.
\end{minted}
% \end{verbatim}

It is because of quotients that we used equivalence relations to define the
components of group structure. If we were to use the regular Leibniz equality,
this would make it very difficult to say that a quotient group is another
group. But by having the definition of a group depend upon an arbitrary
equivalence relation, we enable our theory to state that a quotient group is
simply a group with under a different equivalence. We didn't loose much as we
could still rely upon Coq's setoid rewrite tactics. Setoids are types equipped
with an equivalence relation.

Moving onwards, we then defined structures for rings, which have two binary
operations: a commutative plus \(+\) and a non-commutative times \(\cdot\), and
structures for commutative rings, where multiplication is commutative and has
an identity. Here we define \emph{ideals} which are normal subgroups under
addition and are absorbing with regards to multiplication, i.e. \(r a\) is in
the ideal whenever \(a\) is in the ideal and \(r\) is any element of the ring.
Again, we defined quotient rings but found them unnecessary in the end. Note
that should an ideal \(I\) contain a unit, an element \(x\) with a
multiplicative inverse \(x^{-1}\), then \(I\) would be the entire ring by the
absorbing property: given any element \(a\) of the ring, then \(a x^{-1}\) is
also in the ring, and \(a x^{-1} x = a 1 = a\in I\). Maximal ideals were
defined next. These are ideals that are proper subsets of the ring while having
no larger ideal except for the ring itself. Below is the definition in Coq,
which uses \texttt{P} as the predicate for the ideal.
% \begin{verbatim}
\begin{minted}{coq}
  Definition maximal_ideal :=
    exists (r: Carrier), (not (P r) /\
      forall (Q: Carrier -> Prop)
          (Q_proper: Proper (equiv ==> iff) Q)
          (Q_ideal: Ideal add zero minus mul Q),
        (forall (r: Carrier), P r -> Q r) ->
        (forall (r: Carrier), Q r) \/
          (forall (r: Carrier), Q r -> P r)).
\end{minted}
% \end{verbatim}
Almost by definition, maximal ideals can contain no units, otherwise they would
fail to be a proper subset of the ring. We could then define a local ring,
which is a ring with a single maximal ideal.
% \begin{verbatim}
\begin{minted}{coq}
  Definition local_ring :=
    exists (P: Carrier -> Prop)
        (P_proper: Proper (equiv ==> iff) P)
        (P_ideal: Ideal add zero minus mul P),
      maximal_ideal P /\
      (forall (Q: Carrier -> Prop)
          (Q_proper: Proper (equiv ==> iff) Q)
          (Q_ideal: Ideal add zero minus mul Q),
        maximal_ideal Q -> forall (r: Carrier), P r <-> Q r).
\end{minted}
% \end{verbatim}

@@

\section{Formally Proving Nakayama's Lemma}

>>>>

The set of integers also forms a \emph{ring} due to its properties of multiplication.  
It is closed under this binary operation, which satisfies associativity, and the distributive law governing the compatibility of addition and multiplication holds. 
Even more, the integers form a \emph{commutative ring} 
 since $n \cdot m = m \cdot n$ for all integers $n$ and $m$.  We require commutative rings to contain a \emph{multiplicative identity}, and $1 \in \mathbb{Z}$ is such an element since 
   $n \in \mathbb{Z}$ one has $n \cdot 1 = 1 \cdot n = n$.

@@

The ring of polynomials over a field $F$ in a variable $x$ is not local; in fact, given any irreducible polynomial $f(x)$, the set of its multiples is a maximal ideal of the polynomial ring $F[x]$.
On the other hand, the set of all formal power series  in $x$ over $F$ is a local ring; its maximal ideal consists of the power series with no constant term.

@@

We started by defining a semigroup class, which declares a binary operation to
be associative. From here, we build up through monoids, which introduce
identities, to groups, which introduce inverses. Note the double equals ``$==$'' in
these definitions is notation for an arbitrary equivalence relation over the
group's carrier set, which acts as equality.

\begin{minted}{coq}
  Infix "==" := equiv (at level 60, no associativity).
  Class Semigroup := {
    semigroup_assoc:
      forall (a b c: Carrier),
        a <o> b <o> c == a <o> (b <o> c);
  }.
  Class Monoid := {
    monoid_semigroup :> Semigroup equiv op;
    monoid_ident_l:
      forall (a: Carrier), ident <o> a == a;
    monoid_ident_r:
      forall (a: Carrier), a <o> ident == a;
  }.
  Class Group := {
    group_monoid :> Monoid equiv op ident;
    group_inv_l:
      forall (a: Carrier), inv a <o> a == ident;
    group_inv_r:
      forall (a: Carrier), a <o> inv a == ident;
  }.
\end{minted}
The lines of the form, e.g.,  ``\verb|monoid_semigroup :> Semigroup equiv op;|'' simply coerce
the monoid typeclass into a semigroup.

%While we found later on that we did not need quotients, 
While in the end, our formal proof did not call upon quotients of algebraic structures, quotient rings and quotient modules are fundamental to commutative algebra, and one can use them to construct alternate proofs of Nakayama's lemma.
It is worth pointing out that we have formalized quotients of algebraic objects  in Coq using typeclasses, which appear to work rather nicely.

An algebraic quotient is, roughly, the set of equivalence classes of an algebraic structure with respect to an equivalence relation on its elements, for which the set of equivalence classes inherit the same kind of algebraic structure. 
For example, consider the quotient of a group modulo a subgroup, i.e. a subset of elements of the group that it itself a group under the group operations. 
Under equivalence relation on the group, every element of the subgroup must be in the same equivalence class as the identity.  
With \texttt{P} the predicate for the subgroup, there are two ways to
make an equivalence relation from this description.

\begin{minted}{coq}
  Definition left_congru (a b: Carrier) :=
    P (inv a <o> b).
  Definition right_congru (a b: Carrier) :=
    P (a <o> inv b).
\end{minted}
% \end{verbatim}
When these two relations coincide, then we can prove that this common equivalence
relation actually preserves the group structure. Subgroups which have this
property are called \emph{normal subgroups}.
% \begin{verbatim}
\begin{minted}{coq}
  Let normal_subgroup_congru_coincide :=
    forall (a b: Carrier),
      left_congru op inv P a b <->
      right_congru op inv P a b.

  Theorem quotient_normal_subgroup_group:
    normal_subgroup_congru_coincide ->
    Group (left_congru op inv P) op ident inv.
\end{minted}

The importance of quotients in commutative algebra motivates our use of equivalence relations to define the
components of a group structure. 
If one were to instead use the regular Leibniz equality,
it would be very difficult to identify a quotient group with another
group. 
However, by  defining a group in terms of an arbitrary equivalence relation, we enable our theory to state that a quotient group is
simply a group under a different equivalence. 
Not much is lost,  as Coq's setoid rewrite tactics can still be called upon; 
a setoid is a type equipped
with an equivalence relation.

Moving onward, we formalized the structure of a ring, which has two binary
operations:
addition, which must be commutative, and multiplication, which need not be commutative in general.
Next we defined the structure of a commutative ring, further requiring commutativity of multiplication, as well as a multiplicative identity. 
At this point,, we define {ideals} of commutative rings as normal subgroups of the ring under
addition that satisfy the absorption property under multiplication, i.e., \(r a\) is in
the ideal for every element \(a\) of the ideal, and every element \(r\) of the commutative ring.

Again, we defined quotient rings but found them unnecessary in the end. Note
that should an ideal \(I\) contain a unit, an element \(x\) with a
multiplicative inverse \(x^{-1}\), then \(I\) would be the entire ring by the
absorbing property: given any element \(a\) of the ring, then \(a x^{-1}\) is
also in the ring, and \(a x^{-1} x = a 1 = a\in I\). Maximal ideals were
defined next. These are ideals that are proper subsets of the ring while having
no larger ideal except for the ring itself. Below is the definition in Coq,
which uses \texttt{P} as the predicate for the ideal.
% \begin{verbatim}
\begin{minted}{coq}
  Definition maximal_ideal :=
    exists (r: Carrier), (not (P r) /\
      forall (Q: Carrier -> Prop)
          (Q_proper: Proper (equiv ==> iff) Q)
          (Q_ideal: Ideal add zero minus mul Q),
        (forall (r: Carrier), P r -> Q r) ->
        (forall (r: Carrier), Q r) \/
          (forall (r: Carrier), Q r -> P r)).
\end{minted}
% \end{verbatim}
Almost by definition, maximal ideals can contain no units, otherwise they would
fail to be a proper subset of the ring. We could then define a local ring,
which is a ring with a single maximal ideal.
% \begin{verbatim}
\begin{minted}{coq}
  Definition local_ring :=
    exists (P: Carrier -> Prop)
        (P_proper: Proper (equiv ==> iff) P)
        (P_ideal: Ideal add zero minus mul P),
      maximal_ideal P /\
      (forall (Q: Carrier -> Prop)
          (Q_proper: Proper (equiv ==> iff) Q)
          (Q_ideal: Ideal add zero minus mul Q),
        maximal_ideal Q -> forall (r: Carrier), P r <-> Q r).
\end{minted}
% \end{verbatim}

@@

\section{Formal Proof of Nakayama's Lemma}
